Fusing Filters with Integer Linear Programming

### 2014-05-11 2014betterfusionforfilters/Main.tex
Rewritten intro, typo in the abstract
bench results

\begin{abstract}
In this paper, we present an array fusion technique for functional programs, which is able to handle multi-loop programs containing size altering operations like filter. Recent work on array fusion shows how to extract data-flow graphs from functional programs and compile them into efficient imperative loops.  However, the compilation process only handles graphs that can be compiled into \emph{single} loops, and most programs requiremore. Other work demonstrated how use Integer Linear Programming (ILP) to \emph{cluster} the operators in a general data-flow graph into subgraphs that can be individually fused, but until now this approach did not handle operations like filter, which produce arrays of a different size than their input. The technique we introduce in this paper combines the advantages of both, by  extending the existing ILP approach with support for size changing operators, using an external ILP solver to find good clusterings.

@@

\section*{Acknowledgements}
Many thanks are due to
Robert Clifton-Everest,
Kai Engelhardt,
Bill Kroon,
Frederik Madsen,
Abdallah Saffidine,
and Jingling Xue
for enlightening discussions relating to this work.

>>>>

\begin{abstract}
In this paper, we present an array fusion technique for functional programs, which is able to handle multi-loop programs containing size altering operations like filter. Recent work on array fusion shows how to extract data-flow graphs from functional programs and compile them into efficient imperative loops.  However, the compilation process only handles graphs that can be compiled into \emph{single} loops, and most programs require more. Other work demonstrated how use Integer Linear Programming (ILP) to \emph{cluster} the operators in a general data-flow graph into subgraphs that can be individually fused, but until now this approach did not handle operations like filter, which produce arrays of a different size than their input. The technique we introduce in this paper combines the advantages of both, by  extending the existing ILP approach with support for size changing operators, using an external ILP solver to find good clusterings.

@@

\section*{Acknowledgements}
Many thanks are due to
Robert Clifton-Everest,
Kai Engelhardt,
Bill Kroon,
Frederik Madsen,
Abdallah Saffidine,
Carter Schonwald,
and Jingling Xue
for enlightening discussions relating to this work.

### 2014-05-11 2014betterfusionforfilters/section/05-Benchmarks.tex
bench results

\begin{tabular}{lrrr}
                &   Unfused         & Stream fusion   & \textbf{ ILP               }\\
Normalise2      &   14.0s           & 12.3s           & \textbf{ 11.4s             }\\
Closest pairs   &   25.6s           & 25.1s           & \textbf{ 22.4s             }\\
QuadTree        &   x               & x               & \textbf{ x                 }\\
\end{tabular}

Notes:
\begin{itemize}
\item
Normalise2 tested with around 1gb input data. This should be small enough that all intermediates fit in memory, but large enough not to fit in cache.
\item
ClosestPoints only tested with 80mb data.
\item
ILP solutions were created by hand-fusing based on the clustering of the implementation. In the future, this will be integrated to use data flow fusion.
\item
Benchmark programs are available at \url{https://github.com/amosr/clustering/tree/master/benches/code}
\end{itemize}

>>>>

\begin{figure*}
$$\begin{array}{c}

\begin{tabular}{lrrrrrrrr}
                &   \multicolumn{2}{c}{Unfused}         & \multicolumn{2}{c}{Stream}
                & \multicolumn{2}{c}{Megiddo} &\multicolumn{2}{c}{\textbf{Ours}} \\
                & Time & Loops   & Time & Loops      & Time & Loops & Time & Loops   \\
\hline
Normalise2      & 1.88s & 5      & 1.64s & 4          & 1.82s & 3  & \textbf{1.59s} & \textbf{2}\\
Closest pair    & 3.83s & 6      & 3.33s & 5          & 2.92s & 3  & \textbf{2.92s} & \textbf{3}\\
QuadTree        & 5.22s & 8      & 5.22s & 8          & 4.72s & 2  & \textbf{4.72s} & \textbf{2}\\
\end{tabular}

\end{array}$$
\caption{Benchmark results}
\label{f:BenchResults}
\end{figure*}

Notes:
\begin{itemize}
% \item
% Normalise2 tested with around 1gb input data. This should be small enough that all intermediates fit in memory, but large enough not to fit in cache.
% \item
% ClosestPoints only tested with 80mb data.
\item
In some cases, the clusterings were the same for different methods. In this case, the results are the same. 
\item
The stream fusion benchmarks use the clustering that stream fusion \emph{would} use, but hand-fused and written in C.
\item
Programs were run five times with the same input data, and the fastest run was used.
\item
ILP solutions were created by hand-fusing based on the clustering of the implementation. In the future, this will be integrated to use data flow fusion.
\item
Benchmark programs are available at \url{https://github.com/amosr/papers/tree/master/2014betterfusionforfilters/benches}
\end{itemize}
